---
title: "First Year Exam"
author: "Sneha Patel A12125893"
format: html
---

# Obtaining and opening the data 
First, go to https://data.chhs.ca.gov/dataset/covid-19-variant-data and download the .csv file titled "Covid-19 Variant Data". You can move the file to the same directory as your Quarto document for ease of access. 
Next, we will read the file, and check the first few lines of its contents. 
```{r}
c19 <- read.csv("covid19_variants.csv")

head(c19)
```

# Visualizing the data
We want to graph the data to show the percentage of each variant in all specimens sequenced by month. For this, we will need ggplot2, lubridate, and dplyr 
```{r}
library(ggplot2)
library(lubridate) 
library(dplyr)
```

The `date` column of our data is type <character>. We want to change it to <date> using a lubridate function.
```{r}
c19$date <- ymd(c19$date)
head(c19)
```

Let's plot what we have so far!
```{r}
plot1 <- ggplot(data = c19) + aes(x = date, y = percentage, col=variant_name) + geom_line()
plot1
```
We want to remove data that shows the total number, and other variants. We also only want our data up until May 2022. We will do that with a dplyr function.
```{r}
c19_variants <- c19 %>%
  filter((variant_name != "Other") & (variant_name != "Total"))

c19_variants <- c19_variants %>%
  filter(date < "2022-06-01") 
```

Let's plot what we have so far!
```{r}
p <- ggplot(data = c19_variants) + aes(x = date, y = percentage, col=variant_name) + geom_line()
p
```
This looks great, but lets change the title, axis labels, legend label, and the ticks on the x-axis to show us each month. 
```{r}
p <- p + labs(title = "Covid-19 Variants in California", y = "Percentage of sequenced specimen", x = "Date", col="Variant") + scale_x_date(date_breaks = "month", date_labels = "%b %Y") + theme(axis.text.x = element_text(angle=45, hjust=1))
p
```
WOW! Beautiful :)
